Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-02-06 parallel-tractability-J-v9/response.tex

We thank the reviewers for the time and effort invested in the detailed reviews and the valuable suggestions and comments to improve the quality of the paper.  Below we reply to the individual points of the reviewers.

\section{Rebuttal to Review 1}

\begin{quote}
There should, however, be some discussion somewhere in the paper on the
relationship between path length and KB size.   Even if a KB uses a
non-parallelizable ontology language this doesn't matter much if the paths
that cause problems are short compared to the size of the KB.  The
non-parallel aspects of path processing of short paths will not be noticable
with a small number of threads (certainly up to hundreds) and maybe not even
for very many threads (because of general overhead).
\end{quote}

We added the parallel complexity analysis for Algorithm~$\mathsf{A}_{\text{bsc}}$
and Algorithm~$\mathsf{A}_{\text{opt}}$
in Section~3.3 and Section~3.4 respectively. 
Based on the analysis and Theorem~1, we have that the parallel complexity
of Algorithm~$\mathsf{A}_{\text{bsc}}$ depends only on the graph depths,
while the parallel complexity of Algorithm~$\mathsf{A}_{\text{opt}}$
depends on both of the graph depths and the sizes of inputs.\\

We added the discussion of the graph depths for real ontologies in Section~6.3.
From the evaluation, we observed that the graph depths were far less than
the sizes of the input ontologies. Thus, the reasoning times of the test ontologies
were mainly determined by the input sizes.

@@

\begin{quote}
It would be much better if the ontology names corresponded to the example
numbers, i.e., the ontology in Example 6 was Oex6.
\end{quote}

We made the ontology subscripts correspond to the example numbers
for Example~5, Example~6, Example~7 and Example~9.

\begin{quote}
The notion of a simple concept is very much tied to the algorithm developed
in the paper.  How often do actual ontologies have constructs that could
make them intractable that are rended tractable by having simple concepts?
If there are none, then this class is uninteresting, particularly as the
paper makes such strong claims about worrying about particular ontologies.
This is different from an ontology only containing simple concepts.
\end{quote}

We investigated many real ontologies from different data source. These ontologies cover
several domains and only contain  simple concepts (see Section~6.1). \\

In this work, we focused on exploring the properties
that make ontologies tractable in parallel. Thus, we did not study the problem of making ontologies
tractable in parallel by having simple concepts and its influence. We will study this problem
in the future work.

@@

\begin{quote}
The paper does a bad job of describing why materialization in RDFS is not
parallizable.  RDFS KBs have an infinite number of consequences, so it is
certainly not possible to materialize all of them quickly.  There are,
however, well known tricks that can be used to generate a finite
representation of this infinite set of consequences.  However, even this
cannot be done quickly in parallel.  In fact, even RDFS entailment cannot be
done quickly in parallel, because RDFS allows for a kind of twisted
interaction (which I think it similar to twisted paths in the paper) when
computing subproperty relationships.  Of course, having subproperties of
rdfs:subPropertyOf is not a usual thing to do in RDFS.  The paper needs to
be much more clear on its relationship to RDFS.
\end{quote}

We gave an example that RDFS entailment may lead to the situation of path twisting in Section~4.4 of the revision.
We also discussed in what cases RDFS has parallel tractability of entailment.

\begin{quote}
The paper mentions YAGO early on but does not indicate why YAGO admits good
performance for parallel reasoning.  YAGO uses a very simple ontology
language - RDFS plus two simple extensions.  It is thus a prime candidate
for close examination to show how practical results differ from worst-case
theoretical ones.  However, the paper does not have any examination of
reasoning on YAGO.  Such an examination is very much called for.  Similarly
an examination of the performance of other large RDFS-based KBs, such as
DBpedia, would be very useful.   If YAGO is not benchmarked then it needs to
be dropped entirely.   Similarly for LUBM.
\end{quote}

We extended the discussion for YAGO and LUBM in Section~6.1. We added the evaluation
for these two kinds of ontologies in Section~6.2.

@@

\begin{quote}
The benchmarking part of the paper is very limited and extremely hard to
follow.    Only eight ontologies were benchmarked.  Why were these eight
chosen?  Only considering at most eight threads is much too constricting.
\end{quote}

The eight ontologies are selected since they satisfy the properties for parallel tractability.
We evaluated them to show the difference between the ontologies of tractable and intractable
in parallel.

@@

\begin{quote}
More detail is needed on how the KBs were constructed.  Chain length is a
very important aspect of the KBs, but nothing is said of how the instances
in the KB are connected together.  It is not stated which ontologies belong
to which DL.
\end{quote}

In order to make the experiments close to real scenes, we did not
construct these eight ontologies for some purpose, e.g., to
make the graph depth deeper. Thus, these ontologies acted like common ones.

@@

\begin{quote}
Figure 5 is very misleading because it uses different scales for the two
systems.   There are several immediate questions that arise from this Figure
that are poorly handled in the paper.  First, why does going from one thread
to two make such a difference for ParallelDHL?  Without a convincing
explanation for this surprising aspect of the benchmarking it is hard to
trust any of the rest of the benchmarking.  The other surprising aspect of
Figure 5 is that RDFox speeds up so little.  There is a short explanation of
why this is so, but I would have liked to see a longer one.
\end{quote}

We checked that ParallelDHL costed a large amount of time for computing
the relation $S_{\textit{\!\tiny rch}}$ when only one thread being allocated.
When more than one thread is allocated, ParallelDHL has a better efficiency.
This also resulted in the difference of the speedup between RDFox and ParallelDHL.

@@

\begin{quote}
The computation of the speedup numbers is wholely unexplained.  Why should
this number have any importance?
\end{quote}

We used the speedup numbers to show the distinguishes of reasoning performance between the ontologies in $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ and
$\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$.

@@

\begin{quote}
There is a claim that the benchmarking distinguishes between the Ddhlo and
D-dhlo.  However, Figure 5 doesn't really show this well, if at all.   For
example, why should processing slow down with more threads at all?  Further,
it appears that even for several of the more well-behaved ontologies that there
is very little speedup between 6 and 8 processors.
\end{quote}

We did not
construct the eight ontologies for some purpose.
Thus, these ontologies acted like common ones.
From Figure~5, the distinguishes between the ontologies in
$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ and
$\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$ were not shown
for all the test ontologies.

@@

\begin{quote}
The benchmarking should be redone with more ontologies and with more
threads.  As well, there needs to be better analysis of where parallelism is
failing.  For example, if chains are short, then there should be no problem
in achieving near-perfect thread utilization, ignoring memory contention
problems.
\end{quote}

We added the detailed analysis for LUBM and YAGO and the influence of
graph depths in the revision.

@@

We gave detailed information about the above assumption before Algorithm~$\mathsf{A}_{bsc}$.

\begin{quote}
The computation of the rch relation $S_{\text{rch}}$ ($\dag$) looks problematic, as it requires ``H has been added to G''. If it were the case, then in Example~4 the $S_{\text{rch}}$ would always be empty, as none of the rule heads $H$ would be added to $G$. Again, the complexity analysis of $A_{\text{opt}}$ is unclear and possibly flawed. Step~2 uses an NC algorithm to compute the transitive closure of $S_{\text{rch}}$, which runs in poly-logarithmic time w.r.t.\ the size of $S_{\text{rch}}$ not necessarily so w.r.t.\ the size of $I$. What is the size of $S_{\text{rch}}$ w.r.t.\ that of $I$? Similarly as above, it involves updating $G$, which depends on the sizes of both $G$ and $S_{\text{rch}}$. Would it be still in poly-logarithmic time?
\end{quote}

The requirement ``H has been added to G'' in the description of computing the rch relation $S_{\text{rch}}$ ($\dag$) is indeed
incorrect. It should be ``H has not been added to G''. We fixed this error.\\

We rewrote the paragraph under Lemma~2 to give a more detailed
analysis of the parallel complexity of Algorithm~$\mathsf{A}_{opt}$.
In Step~2 of Algorithm~$\mathsf{A}_{opt}$, an NC algorithm is used to compute the transitive closure of $S_{\text{rch}}$.
It can be checked that the scale of $S_{\text{rch}}$ is polynomial in the size of inputs.
The total computing time is still in poly-logarithmic time (see the detailed analysis in the revision).

@@

\begin{quote}
Finally, the algorithms assume each processor handling a rule instantiation. I wonder how it is implemented in ParallelDHL.
\end{quote}

We added the discussion of how each processor handles rule instantiations at the beginning of Section~6.2.

### 2018-02-06 parallel-tractability-J-v9/parallelTractabilityDatalog.tex
added the experiments of lubm and yago
added the response for the two reviewers

\begin{theorem}\label{theorem:a1}
For any datalog program $P$, $P\in\mathcal{D}_{\mathsf{A}_{\text{bsc}}^{\psi}}$ iff $P$ has a
materialization graph whose depth is upper-bounded by $\psi(|P|)$.
\end{theorem}

@@

\subsection{Optimizing Algorithm~$\mathsf{A}_{\text{bsc}}$ via Single-Way Derivability}

>>>>

\begin{theorem}\label{theorem:a1}
For any datalog program $P=\langle R, \textbf{I}\rangle$, $P\in\mathcal{D}_{\mathsf{A}_{\text{bsc}}^{\psi}}$ iff $P$ has a
materialization graph whose depth is upper-bounded by $\psi(|\textbf{I}|)$.
\end{theorem}

@@

\subsection{Optimizing Algorithm~$\mathsf{A}_{\text{bsc}}$ via Single-Way Derivability}
\label{sec:opt}
